Scheduling and Amortization for MPC

### 2021-11-22 paper_SIMD.tex
Fixed typos.
added my comments/questions

\begin{itemize}
\item same-level (forward) $X\rightarrow Y$ where $X = ...$ and $Y = ...$ are in the same loop nest $i_1,i_2,...i_k$. E.g., $\texttt{d} = \mathit{SUM}(\texttt{S}[i,j], \texttt{C}[j])$ to $\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$ in Biometric is a same-level edge. 
\item outer-to-inner (forward) $X\rightarrow Y$ where $X$ is in an outer loop nest, $i_1,i_2,...i_j$, and $Y$ is in an inner one, $i_1,i_2,...i_j,...,i_k$.
\item inner-to-outer (forward) $X\rightarrow Y$ where $X$ is in an inner loop nest, $i_1,i_2,...i_j,...,i_k$, and $Y$ is in an outer one, $i_1,i_2,...i_j$. E.g. $\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_1,\texttt{min}_0)$ is an inner-to-outer edge.
\item same-level back-edge $X\rightarrow Y$. $Y$ is a phi-node in the header of the loop and $X$ is a definition of the variable in the loop body. E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
\item outer-to-inner back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in an immediately nested loop of $Y$'s loop. 
 \end{itemize}

@@

\section{Basic Vectorization}

@@

@@

\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i_1,...,i_k$}
\FOR {each $Y_i$}
\STATE {case def-use edge $Y_i \rightarrow X$ of} \\
\STATE \hspace{0.25cm} {same-level $\rightarrow$ add $Y_i = Y_i$} \\
\STATE \hspace{0.25cm} {outer-to-inner $\rightarrow$ add $Y'_i[i_1,...,i_k] = \mathit{raise\_dim}(Y_i)$} 
\STATE \hspace{0.25cm} {inner-to-outer $\rightarrow$ add $Y'_i[i_1,...,i_k] = \mathit{drop\_dim}(Y_i)$} 
\ENDFOR

\COMMENT {Optimistically vectorize all. $\stackrel{\rightarrow}{i}$ means vectorized dimension.}
\STATE {change to $X[\stackrel{\rightarrow}{i_1},...\stackrel{\rightarrow}{i_k}] = \mathit{Op}(Y'_1[i_1,...,i_k],Y'_2[i_1,...,i_k])$ }
\ENDFOR

@@

@@

@@

@@

Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$ and $\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$.
The closure of the first is 
\[
\begin{array}{l}
\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1), \\ 
\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N]) 
\end{array}
\]
and the closure of the second one is
\[
\begin{array}{l}
\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000) \\
\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_1[\stackrel{\rightarrow}{i}])
\end{array}
\]
Since the two closures intersect, we merge them into a single loop. We can now rewrite those statements, getting rid of the vectorization on $i$ in second FOR loop. Notably, the SUB and MUL computations are fully vectorizable and the ADD computation is vectorizable across the $i$-dimension.

>>>>

\begin{itemize}
\item same-level (forward) $X\rightarrow Y$ where $X = ...$ and $Y = ...$ are in the same loop nest $i_1,i_2,...i_k$. E.g., $\texttt{d} = \mathit{SUM}(\texttt{S}[i,j], \texttt{C}[j])$ to $\texttt{p} = \mathit{MUL}(\texttt{d},\texttt{d})$ in Biometric is a same-level edge. 
\item outer-to-inner (forward) $X\rightarrow Y$ where $X$ is in an outer loop nest, $i_1,i_2,...i_j$, and $Y$ is in an inner one, $i_1,i_2,...i_j,...,i_k$.
\item inner-to-outer (forward) $X\rightarrow Y$ where $X$ is in an inner loop nest, $i_1,i_2,...i_j,...,i_k$, and $Y$ is in an outer one, $i_1,i_2,...i_j$. E.g. $\texttt{sum}_1 = \texttt{ADD}(\texttt{sum}_0,\texttt{p})$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_1,\texttt{min}_0)$ is an inner-to-outer edge.
\item same-level back-edge $X\rightarrow Y$. $Y$ is a phi-node in the header of the loop and $X$ is a definition of the variable in the loop body. E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
\item inner-to-outer back-edge $X\rightarrow Y$: $X$ and $Y$ are both $\phi$-nodes for some variable. The source $X$ is in an immediately nested loop of $Y$'s loop. 
 \end{itemize}

@@

\section{Vectorization}

@@

\subsection{Basic Vectorization}

@@

\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i_1,...,i_k$}
\FOR {each $Y_i$}
\STATE {case def-use edge $Y_i \rightarrow X$ of} \\
\STATE \hspace{0.25cm} {same-level $\rightarrow$ add $Y'_i = Y_i$} \\
\STATE \hspace{0.25cm} {outer-to-inner $\rightarrow$ add $Y'_i[i_1,...,i_k] = \mathit{raise\_dim}(Y_i)$} 
\STATE \hspace{0.25cm} {inner-to-outer $\rightarrow$ add $Y'_i[i_1,...,i_k] = \mathit{drop\_dim}(Y_i)$} 
\ENDFOR

\COMMENT {Optimistically vectorize all. $\stackrel{\rightarrow}{i}$ means vectorized dimension.}
\STATE {change to $X[\stackrel{\rightarrow}{i_1},...\stackrel{\rightarrow}{i_k}] = \mathit{Op}(Y'_1[\stackrel{\rightarrow}{i_1},...,\stackrel{\rightarrow}{i_k}],Y'_2[\stackrel{\rightarrow}{i_1},...,\stackrel{\rightarrow}{i_k}])$ }
\ENDFOR

@@

\ishaq{We need to think of a way to handle 4th case for def-use, when a nested block B def is followed by another, different, nested block B' use. Note: The handwritten notes say we can break it into 2 edges. One idea to this end is to insert a def' node which is a copy of the original def. The def' should be placed in a block that contains both B and B'.}

@@

\ishaq{At least for input variables: when raising dimensions, we should keep some meta data around so we can later write code that could use 1 share if it knows the other N shares are just duplicates of it.}

\ishaq{Naive question: Which index of $\texttt{min}_{1}[\stackrel{\rightarrow}{i}]$ would contain the minimum, how are we reducing it?}

@@

\ishaq{For our rader: Such dependencies will not involve phi nodes only. e.g. $a[i] = a[i] + a[i-1]$}

@@

Next, analysis moves to outer dimension $i$. There are two phi-nodes, $\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$ and $\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000)$.
The closure of the first is 
\[
\begin{array}{l}
\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1), \\ 
\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
\texttt{index}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{index}_0[\stackrel{\rightarrow}{i}],[0,1,...N]) 
\end{array}
\]
and the closure of the second one is
\[
\begin{array}{l}
\texttt{min}_0[i] = \phi(\texttt{min}_1[i-1], 10000) \\
\texttt{c}[\stackrel{\rightarrow}{i}] = \mathit{CMP}(\texttt{sum}'_1[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}]), \\
\texttt{min}_1[\stackrel{\rightarrow}{i}] = \mathit{MUX}(\texttt{c}[\stackrel{\rightarrow}{i}],\texttt{min}_0[\stackrel{\rightarrow}{i}],\texttt{sum}'_1[\stackrel{\rightarrow}{i}])
\end{array}
\]

\ishaq{I don't think I understand closures. Why does the closure for $\texttt{index}_0[i]$ include $\texttt{c}[\stackrel{\rightarrow}{i}]$ but the closure for $\texttt{index}_0[i] = \phi(\texttt{index}_1[i-1], -1)$ not include $\texttt{p}$.}

Since the two closures intersect, we merge them into a single loop. We can now rewrite those statements, getting rid of the vectorization on $i$ in second FOR loop. Notably, the SUB and MUL computations are fully vectorizable and the ADD computation is vectorizable across the $i$-dimension.
